\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\ \large\textbf{ Preparation of Papers in Two-Column Format\\ for the ASP-DAC 2014 (\LaTeX2e version)}

### 2014-01-27 motivation.tex
Updated the paper

\subsection{Motivating Example}
\label{sec:motivating-example}

% system{
%  interface{
%   input signal TOUCH;
%   output signal GREEN_LIGHT;
%   output char channel S;
%   input char channel R;
%  }
\begin{figure}[t!]
% \begin{center}
	\vspace{-10pt}
		\begin{lstlisting}[style=sysj,morekeywords={sustain,send,receive,abort,await,emit,present,trap,pause,exit,delay,suspend}]
{ // Clock-domain 1
while(true) {
  trap(TRAP){
   // Abort if user touches the screen
   abort(TOUCH){
    {sustain GREEN_LIGHT;} //reaction R1
    || // synchronous parallel operator
    {
     //exit after any where from 50.3 to 200.3 ms.
     delay (50.3 .. 200.3 ms); 
     exit(TRAP); 
    } // reaction R2
   } do {send S(1);} // Send 1 if touched 
  } do {send S(0);} // Else Send 0
  pause; // only construct to consume logical time
 }
}
>< // asynchronous operator
{ // Clock-domain 2
  while(true){ receive S; /* do something with S */}
}
\end{lstlisting}
\caption{An example human responsiveness system (HRTCS)}
% }
% \end{center}
\label{fig:1}
\end{figure}

@@

The SystemJ pseudo-code~\cite{amal10} for such a system is shown in
Figure~\ref{fig:1}. We use the SystemJ language in this paper, first of
all, because being a \textit{Globally Asynchronous Locally Synchronous}
(GALS) language it helps us easily describe a larger class of systems
(like HRTCS in Figure~\ref{fig:1}) compared to purely synchronous
languages like Esterel. Moreover, being a super-set of Esterel, SystemJ
allows us to explore the marriage of real-time and logical time, which
is equally applicable to the synchronous sub-set.

The SystemJ program is divided into two clock-domains, the first
clock-domain continuously displays a green light (\texttt{GREEN\_LIGHT}
signal) on the touch screen sensor and waits for the user to respond. If
the user is able to touch the screen within the specified time of 50.3
ms to 200.3 ms a positive response is sent to the second
clock-domain. All the SystemJ programming constructs used to implement
this system are described in~\cite{amal10} and Table~\ref{tab:1} in
Section~\ref{sec:background}. The delay specification, which models the
non-deterministic delay between 50.3 ms and 200.3 ms is not supported in
the current language and cannot be compiled with the current SystemJ
compiler. We introduce real-time \texttt{delay} statements in SystemJ.

>>>>

\subsection{Motivation via examples of real-time systems programmed in a
reactive language -- SystemJ}
\label{sec:motivating-example}

% system{
%  interface{
%   input signal TOUCH;
%   output signal GREEN_LIGHT;
%   output char channel S;
%   input char channel R;
%  }

Real-time operating systems~\cite{barry2009using} usually provide two
types of mechanisms for introducing real-time in the developed program:
(1) the ability to perform a timeout and (2) the ability to run tasks
periodically with some real-time period $T$. We would like to provide
similar mechanisms in reactive languages for real-time program
development. In fact, we provide \textit{exact} and \textit{non exact}
real-time control mechanisms -- a generalization of the two real-time
mechanisms introduced in real-time operating systems.

\subsubsection{Programming using the non exact real-time control
  construct the \textrm{\texttt{wait\_inbetween}} statement}
\label{sec:progr-using-non}

\begin{figure}[t!]
	\vspace{-10pt}
        \begin{SubFloat}{\label{delaycode:a}A human responsiveness
            system (HRTCS)}
        \begin{lstlisting}[style=sysj,morekeywords={sustain,send,receive,abort,await,emit,present,trap,pause,exit,wait_inbetween,suspend}]
{ // Clock-domain 1
while(true) {
  trap(TRAP){
   // Abort if user touches the screen
   abort(TOUCH){
    {sustain GREEN_LIGHT;} //reaction R1
    || // synchronous parallel operator
    {
     //exit after any where from 50.3 to 200.3 ms.
     wait_inbetween (50.3 .. 200.3 ms); 
     exit(TRAP); 
    } // reaction R2
   } do {send S(1);} // Send 1 if touched 
  } do {send S(0);} // Else Send 0
  pause; // only construct to consume logical time
 }
}
>< // asynchronous operator
{ // Clock-domain 2
  while(true){ receive S; /* do something with S */}
}
\end{lstlisting}
\end{SubFloat}
\begin{SubFloat}{\label{delaycode:b}Timing diagram -- the X-axis shows
    the execution of the \texttt{wait\_inbetween} statement. The Y-axis
    is the time elapsed.}
  % \begin{figure}[h!]
    \centering
    \includegraphics[scale=0.7]{FF1}
  % \end{figure}
\end{SubFloat}
\caption{Programming a human responsiveness system (HRTCS) with
  \texttt{wait\_inbetween}}
% }
% \end{center}
\label{fig:1}
\end{figure}

@@

The SystemJ pseudo-code~\cite{amal10} for such a system, along with the
expected timing behavior of the \texttt{wait\_inbetween} statement, is
shown in Figure~\ref{fig:1}. We use the SystemJ language in this paper,
first of all, because being a \textit{Globally Asynchronous Locally
  Synchronous} (GALS) language it helps us easily describe a larger
class of systems (like HRTCS in Figure~\ref{fig:1}) compared to purely
synchronous languages like Esterel. Moreover, being a super-set of
Esterel, SystemJ allows us to explore the marriage of real-time and
logical time, which is equally applicable to the synchronous sub-set.

The SystemJ program is divided into two clock-domains, the first
clock-domain continuously displays a green light (\texttt{GREEN\_LIGHT}
signal) on the touch screen sensor and waits for the user to respond. If
the user is able to touch the screen within the specified time of 50.3
ms to 200.3 ms a positive response is sent to the second
clock-domain. All the SystemJ programming constructs used to implement
this system are described in~\cite{amal10} and Table~\ref{tab:1}. The
\texttt{wait\_inbetween} specification, which models the non exact delay
between 50.3 ms and 200.3 ms is not supported in the current language
and cannot be compiled with the current SystemJ compiler. We introduce
real-time \texttt{wait\_inbetween} statements in SystemJ. Such non exact
real-time mechanisms are useful in developing systems where the
real-time delay is not know or should not be known a priori.

\subsubsection{Programming a timeout using the non exact real-time
  construct the \texttt{wait\_atleast} statement}
\label{sec:progr-time-using}

\begin{figure}[b!]
	\centering
	\vspace{-10pt}
        \begin{SubFloat}{\label{dd:a}Timeout waiting for the input
            signal \texttt{DoorOpened}}
        \begin{lstlisting}[style=sysj,morekeywords={abort,await,emit,present,trap,pause,exit,wait_atleast,suspend}]
trap(T){
  abort(DoorOpened){
    wait_atleast(10000 ms);
    exit(T);  
  }
}
\end{lstlisting}
\end{SubFloat}
\begin{SubFloat}{\label{dd:b}Timing diagram -- the X-axis shows the
    execution of the \texttt{wait\_atleast} statement. The Y-axis is the
    time elapsed.}
\includegraphics[scale=0.7]{FF2}
\end{SubFloat}
\caption{Programming a timeout with \texttt{wait\_atleast}}
\label{dd}
\end{figure}

There are many instances when one would like to wait on an input from
environment for only a specified amount of time. If the signal is not
received then a timeout is generated so that the program can make
progress. Such a timeout programmed in the SystemJ pseudo-code along
with its timing behavior is shown in Figure~\ref{dd}. The SystemJ
program is waiting on an input signal, \texttt{DoorOpened}, from the
environment for 10 seconds and makes progress irrespective of the
reception of this signal after 10 seconds.

The timing diagram, Figure~\ref{dd:b}, shows that the
\texttt{wait\_atleast} statement waits for a \textit{at least} 10
seconds, but unlike \texttt{wait\_inbetween} there is no upper
bound. The statement following the \texttt{wait\_atleast} statement is
executed after a finite countable, but unbounded delay. Such a statement
is useful for programming low priority tasks.

\subsubsection{Programming periodic tasks using the exact real-time
  construct the \texttt{wait\_exactly} statement}
\label{sec:progr-using-exact}

The previous wait statements are non exact, i.e., the execution of the
statement following the wait statement cannot be controlled exactly. The
final variant we introduce allows doing exactly that. We call it
\texttt{wait\_exactly} and this statement can be used to program exact
timeouts, periodic tasks, etc.

\begin{figure}[t!]
  \centering
	\vspace{-10pt}
        \begin{SubFloat}{\label{pp:a}Periodically emitting signal S}
        \begin{lstlisting}[style=sysj,morekeywords={emit,trap,pause,exit,wait_exactly}]
while(true) { 
 wait_exactly (1 ms); 
 emit S; 
 //do something 
}
\end{lstlisting}
\end{SubFloat}
\begin{SubFloat}{\label{pp:b}Timing diagram -- the X-axis shows the
    execution of the \texttt{wait\_exactly} statement. The Y-axis is the
    time elapsed.}
  \includegraphics[scale=0.7]{FF3}
\end{SubFloat}
  \caption{Programming a periodic task with wait\_exactly}
  \label{fig:periodic}
  \vspace{-10pt}
\end{figure}

One such example; periodic emission of a signal to the environment is
shown in Figure~\ref{fig:periodic}. As before upon execution of
\texttt{wait\_exactly (1 ms)} statement, the time starts elapsing. But,
unlike in previous cases the following statement needs to be executed
once 1 ms have passed, as shown in Figure~\ref{pp:b}.

### 2014-01-27 related_work.tex
Updated the paper

A number of studies have been performed when it comes to integrating
real-time directly as a programming construct in GALS (and its subset
synchronous) reactive languages. The most prominent works in this area
are by Shyamsundar~\cite{rsh94} and Bourke et
al.~\cite{Bourke2009a}. Shyamsundar incorporates real-time using
external timers in \textit{Communicating Reactive Processes} (CRP),
which like SystemJ is an extension of synchronous language Esterel to
asynchorny. But, as mentioned in Section~\ref{sec:inter-preempt-delays},
external timers do not interact well with preemption constructs in these
languages. Bourke et al. introduce real-time as first class constructs
in the Esterel language. They like us provide real-time \texttt{delay}
as first class Esterel programming constructs, and translate them into
Esterel kernel constructs. But, unlike us they do not translate delays
into \texttt{pause} constructs directly. Instead, logical ticks are
generated by using abstract notion of \texttt{event} and \texttt{sample}
platform dependent timers. This notion makes it a complex and inflexible
solution since the number of platform timers with certain resolutions
need to be determined and present on the system for the solution to be
realizable. Moreover, non-deterministic real-time delays and integration
with non-maximal parallelism is not studied at all. Our solution does
not require external timers (like~\cite{rsh94}) or logical notion of
timers (like~\cite{Bourke2009a}). In our solution we calculate the delay
in logical ticks \textit{after} allocation and scheduling, it gives the
compiler developers the chance to optimize for many different criteria
such as computation time, energy, power, etc, without worrying about or
violating the real-time delay specification.

Other works such as Quartz~\cite{glog02} and Bertin et
al.~\cite{Bertin:2000:TVR:1947412.1947439}, also incorporate delay
statements in Esterel using \texttt{pause} and special \textit{pragmas}
(annotations), respectively.  The Quartz solution is targeted at
studying timing properties using model checkers rather than
implementation, like us. The solution introduced by Bertin et
al.~\cite{Bertin:2000:TVR:1947412.1947439} translates real-time pragmas
into Timed automata~\cite{alur94} for model-checking real-time
deadlines. Our solution on the other hand removes all continuous time
elements and produces single discrete time model thereby making the
solution amenable to existing functional and real-time verification
tools developed for reactive languages and more in the spirit of logical
time. Finally, a major difference between Bertin et al. and our approach
is that their approach requires programmers to annotate the time taken
by `C' code fragments manually, this is extremely error prone, we use
static low-level program analysis to find out such times automatically.

>>>>

A number of studies have been performed when it comes to integrating
real-time directly as a programming construct in GALS (and its subset
synchronous) reactive languages. The most prominent works in this area
are by Shyamsundar~\cite{rsh94} and Bourke et
al.~\cite{Bourke2009a}. Shyamsundar incorporates real-time using
external timers in \textit{Communicating Reactive Processes} (CRP),
which like SystemJ is an extension of synchronous language Esterel to
asynchorny. But, as mentioned in Section~\ref{sec:inter-timers-react}
and~\ref{sec:resolution-real-time} external timers do not interact well
with preemption constructs in these languages. Bourke et al. introduce
real-time as first class constructs in the Esterel language. They like
us provide real-time wait statements called \texttt{delay} as first
class Esterel programming constructs, and translate them into Esterel
kernel constructs. But, unlike us they do not translate wait statements
into \texttt{pause} constructs directly. Instead, logical ticks are
generated by using abstract notion of \texttt{event} and \texttt{sample}
platform dependent timers. This notion makes it a complex and inflexible
solution since the number of platform timers with certain resolutions
need to be determined and present on the system for the solution to be
realizable. Moreover, non exact real-time waits and integration with
non-maximal parallelism is not studied at all. Our solution does not
require external timers (like~\cite{rsh94}) or logical notion of timers
(like~\cite{Bourke2009a}). In our solution we calculate the real-time
waiting time in logical ticks \textit{after} allocation and scheduling,
it gives the compiler developers the chance to optimize for many
different criteria such as computation time, energy, power, etc, without
worrying about or violating the real-time wait specification.

Other works such as Quartz~\cite{glog02} and Bertin et
al.~\cite{Bertin:2000:TVR:1947412.1947439}, also incorporate real-time
wait statements in Esterel using \texttt{pause} and special
\textit{pragmas} (annotations), respectively.  The Quartz solution is
targeted at studying timing properties using model checkers rather than
implementation, like us. The solution introduced by Bertin et
al.~\cite{Bertin:2000:TVR:1947412.1947439} translates real-time pragmas
into Timed automata~\cite{alur94} for model-checking real-time
deadlines. Our solution on the other hand removes all continuous time
elements and produces single discrete time model thereby making the
solution amenable to existing functional and real-time verification
tools developed for reactive languages and more in the spirit of logical
time. Finally, a major difference between Bertin et al. and our approach
is that their approach requires programmers to annotate the time taken
by `C' code fragments manually, this is extremely error prone, we use
static low-level program analysis to find out such times automatically.
